---
title: "LMF Weights"
author: "Alex Traynor"
date: "6/23/2021"
output: html_document
---
### Setup
Set local library path and load required packages
```{r setup, eval = TRUE}

.libPaths(c("C:/R-4.0.3/library", .libPaths()))

library(aim.analysis)
library(sp)
library(dplyr)
library(stringr)
library(readxl)
library(rgdal)
library(arcgisbinding)
arc.check_product()

`%notin%` <- Negate(`%in%`)

knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Projection to use. Alber's Equal Area is useful for calculating areas
projection <- sp::CRS("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")

# Date of analysis
date <- Sys.Date()

# Confidence level in percent
confidence <- 80

# location where we will write analysis csvs
output_path <- "C:\\Users\\alaurencetraynor\\Documents\\2021\\Analysis\\NRCS\\aim_lmf_weights\\output"

# GDB within the output_path for writing points and wgtcat spatial outputs
output_gdb <- "LMF_Weights.gdb"

# Core filename for all outputs for this reporting unit
output_filename <- "LMF_rangeland"

# The variable in LMF points that contains the unique identifier for each point
lmf_idvar <- "LMF_final_join_PrimaryKey"

# The fates in sample designs which correspond to sampled/observed
observed_fates = c("TS", "Target Sampled", "Sampled")

# The variable containing the LMF segment identity/membership
# This is found in the points and the segment polygons
segment_var <- "segment_id"

# The variable that contains the areas in HECTARES in wgtcats
# This is calculated in ArcMap
wgtcat_area_var = "POLY_AREA"

# The unique ID for poststratification polygons
wgtcats_var <- "stratum"

### The filepath to the folder where we have all the spatial data
path_spatial <- "C:\\Users\\alaurencetraynor\\Documents\\2021\\Analysis\\NRCS\\aim_lmf_weights\\input"

# The geodatabase containing all the spatial files
analysis_gdb <- "aim_lmf_weights.gdb"

### Spatial file names
# name of post-stratification file for this reporting unit
strata <- "strata_BLM_LRR_counties_nlcdrange"

# name of sampled lmf points file in analysis gdb
lmf_file <- "LMF_20142020_segmentjoin_strata"

# variables in points describing the area ratio of lmf segments
lmf_area_var <- "area_ratio"
```

### Read in

#### Strata polygons
```{r}
wgtcat_df <-sf::st_read(dsn = paste(path_spatial, analysis_gdb, sep = "/"),
                        layer = strata,
                        stringsAsFactors = FALSE,
                        type = 3)

wgtcat_df[["wgtcat"]] <-  as.character(wgtcat_df[[wgtcats_var]])
wgtcat_df[["area_hectares"]] <- wgtcat_df[[wgtcat_area_var]]

wgtcat_df <- wgtcat_df[,c("wgtcat","area_hectares")]


# Make Spatial version for wgtcat summary output
wgtcat_spdf <- sf::st_transform(wgtcat_df,
                                crs = projection)

wgtcat_spdf <- sf::st_zm(wgtcat_df,
                            drop = TRUE,
                            what = "ZM") # simplify in case theres Z/M values

wgtcat_spdf <- methods::as(wgtcat_df, "Spatial")

# Non-spatial for weight calcs
sf::st_geometry(wgtcat_df) <- NULL
wgtcats <- as.data.frame(wgtcat_df)
```

#### lmf points
```{r}
### add points
lmf_points <- sf::st_read(dsn = paste(path_spatial, analysis_gdb, sep = "/"),
                           layer = lmf_file,
                           stringsAsFactors = FALSE)
## Standardize projection
lmf_points <- sf::st_transform(lmf_points,
                                crs = projection)

lmf_points <- sf::st_zm(lmf_points,
                            drop = TRUE,
                            what = "ZM") # simplify becuase theres Z/M values

## Make spatial 
lmf_points <- methods::as(lmf_points, "Spatial")

```
###### Sort out points ######
#### Prepare TerrADat points
```{r}
lmf_points@data[["unique_id"]] <- lmf_points@data[[lmf_idvar]]
lmf_points@data[["fate"]] <- observed_fates[1]
lmf_points@data[["wgtcat"]] <- lmf_points@data[[wgtcats_var]]
lmf_points@data[["segment"]] <- lmf_points@data[[segment_var]]
lmf_points@data[["area_ratio"]] <- lmf_points@data[[lmf_area_var]]
lmf_points@data <- lmf_points@data[,c("unique_id", "fate","wgtcat","segment", "area_ratio")]


# Add in coordinates (needed for the analysis step)
lmf_points <-  aim.analysis::add_coords(lmf_points,
                                        xynames = c("x", "y"))

# double check this is character
lmf_points$wgtcat <- as.character(lmf_points$wgtcat)
```

# Filter to those points which have landhealth data
```{r}
list_path <- "C:\\Users\\alaurencetraynor\\Documents\\2021\\Analysis\\NRCS\\aim_lmf_weights\\input\\Copy of revised_LMF_point_counts_NEED_New_Weights_8-30-2021.xlsx"

list <- readxl::read_xlsx(list_path, sheet =  "6707 final pts that need wgts")

lmf_points <- lmf_points[lmf_points$unique_id %in% list$unique_id,]
```

###### Analyze ######
############################################################
#### Calculate weights ####
### Note that this is the contents of aim.analysis::weight_aimlmf() verbatim
### As a function, it was producing errors that it does not as bare code
############################################################

```{r}
# These are being redefined here as though they were arguments to the function
lmf_idvar <- "unique_id"
wgtcat_var <- "wgtcat"
wgtcat_area_var <- "area_hectares"
segments <- NULL
segment_var <- "segment"
verbose <- FALSE 


# Add reporting units
lmf_points[["reporting_unit"]] <- output_filename

# If somehow LMF points aren't in a segment, that's a major problem
if (any(is.na(lmf_points[["segment"]]))) {
  stop(paste("The following LMF points did not spatially intersect any segment polygons:",
             paste(lmf_points[is.na(lmf_points[["segment"]]), "unique_id"], collapse = ", ")))
}

# TODO: Stick a check in here that the segment ID from the polygons
# matches the one derived from the LMF plot ID
# Probably just warn if not?


# Get data frames
if (class(lmf_points) %in% "SpatialPointsDataFrame") {
  lmf_df <- lmf_points@data
} else {
  lmf_df <- lmf_points
}

# NOTE THAT THIS FILTERS OUT ANYTHING FLAGGED AS NOT NEEDED IN THE FATE
# So that'd be unused oversamples or points from the FUTURE that no one would've sampled anyway

# We only have target sampled LMF points available to us, so we don't need to filter them
lmf_df <- lmf_df[, c("unique_id", "fate", "wgtcat", "segment", "area_ratio")]
lmf_df[["aim"]] <- FALSE
lmf_df[["lmf"]] <- TRUE

# Combine them
combined_df <- lmf_df

# There shouldn't be any that don't belong to a wgtcat anymore
combined_df <- combined_df[!is.na(combined_df[["wgtcat"]]), ]

# Add an observed variable for easy reference later
combined_df[["observed"]] <- combined_df[["fate"]] %in% observed_fates

# To make the lookup table, drop any points that fell outside LMF segments
combined_segmentsonly_df <- combined_df[!is.na(combined_df[["segment"]]), ]

# Create a segment relative weight lookup table
segment_relwgt_lut <- do.call(rbind,
                              lapply(X = split(combined_segmentsonly_df, combined_segmentsonly_df[["segment"]]),
                                     FUN = function(X){
                                       
                                       # We also need the count of LMF points with any valid fate, but that's complicated
                                       # We only have the sampled LMF points so we can count those
                                       lmf_sampled_count <- sum(X[["lmf"]])
                                       # To get the number of evaluated but not sampled points:
                                       # The LMF plot keys end in a digit that represents the intended sampling order within a segment
                                       # 1 and 2 are considered base points and were intended to be sampled
                                       # If a sampled LMF plot's plot key ends in 3, that means that one or both of the base points
                                       # were evaluated and rejected rather than sampled, which brings the evaluated LMF plot count
                                       # to three for the segment.
                                       # This just asks if the third point was used
                                       lmf_oversample_used <- any(grepl(X[["unique_id"]][X[["lmf"]]],
                                                                        pattern = "\\D3$"))
                                       
                                       # Likewise, if only one LMF plot was sampled in a segment, that means the other two were
                                       # evalurated and rejected rather than sampled, also bringing the total to three.
                                       # So if there was only one sampled or if the oversample was used, there were three evaluated
                                       if (sum(X[["lmf"]]) == 1 | lmf_oversample_used) {
                                         lmf_count <- 3
                                       } else {
                                         # This will fire only if there sampled count was 2, but better to be safe here
                                         lmf_count <- lmf_sampled_count
                                       }
                                       
                                       
                                       # The relative weight for points falling within a segment is calculated as
                                       # 1 / (number of points)
                                       relative_weight <- (1 / lmf_count)
                                       
                                       output <- data.frame("segment" = X[["segment"]][1],
                                                            "relwgt" = relative_weight,
                                                            stringsAsFactors = FALSE)
                                       return(output)
                                     }))

# Add the relative weights to the combined AIM and LMF points
combined_df <- merge(x = combined_df,
                     y = segment_relwgt_lut,
                     by = "segment",
                     all.x = TRUE)

# adjust relwgt based on area of lmf segment:
combined_df <- combined_df %>% mutate(area_ratio = round(area_ratio, digits = 3)) %>%
                               mutate(relwgt = relwgt * area_ratio)

if (is.null(wgtcat_area_var)) {
  if (class(wgtcats) %in% "SpatialPolygonsDataFrame") {
    wgtcat_df <- aim.analysis::add.area(wgtcats)@data
  } else {
    stop("No name for a variable in wgtcats containing the area in hectares was provided and wgtcats is not a spatial polygons data frame so area can't be calculated")
  }
} else {
  if (!(wgtcat_area_var %in% names(wgtcats))) {
    if (class(wgtcats) %in% "SpatialPolygonsDataFrame") {
      wgtcat_df <- aim.analysis::add.area(wgtcats)@data
    } else {
      stop("The variable ", wgtcat_area_var, " does not appear in wgtcats")
    }
  } else {
    warning("Trusting that the variable ", wgtcat_area_var, " in wgtcats contains the areas in hectares")
    if (class(wgtcats) %in% "SpatialPolygonsDataFrame") {
      wgtcat_df <- wgtcats@data
      wgtcat_df[["AREA.HA"]] <- wgtcat_df[[wgtcat_area_var]]
    } else {
      wgtcat_df <- wgtcats
      wgtcat_df[["AREA.HA"]] <- wgtcat_df[[wgtcat_area_var]]
    }
  }
}


wgtcat_df[["wgtcat"]] <- wgtcat_df[[wgtcat_var]]

# Making sure that these aren't spread out over multiple observations
if (verbose) {
  message("Summarizing wgtcat_df by wgtcat to calculate areas in case the wgtcats are split into multiple observations")
}
wgtcat_df <- dplyr::summarize(dplyr::group_by(wgtcat_df,
                                              wgtcat),
                              "hectares" = sum(AREA.HA))

wgtcat_areas <- setNames(wgtcat_df[["hectares"]], wgtcat_df[["wgtcat"]])

# had a issue here reading wgtcats as integers
wgtcat_df$wgtcat <-  as.character(wgtcat_df$wgtcat)
  
weight_info <- lapply(X = wgtcat_df[["wgtcat"]],
                      wgtcat_areas = wgtcat_areas,
                      points = combined_df,
                      wgtcat_var = wgtcat_var,
                      FUN = function(X, wgtcat_areas, points, wgtcat_var){
                        # Area of this polygon
                        area <- wgtcat_areas[X]
                        
                        # All of the points (AIM and LMF) falling in this polygon
                        points <- points[points[["wgtcat"]] == X, ]
                        
                        # If there are in fact points, do some weight calculations!
                        if (nrow(points) > 0 & any(points[["observed"]])) {
                          # The number of observed AIM points with a relative weight of 1
                          # So, not sharing an LMF segment with any LMF points
                          aim_standalone <- sum(combined_df[["aim"]] & combined_df[["relwgt"]] == 1)
                          # The number of unique segments selected for LMF points
                          lmf_segment_count <- length(unique(points[["segment"]]))
                          
                          # The approximate number of 160 acre segments in this polygon
                          # Obviously, not all of the segments were selected in the first stage
                          # of the LMF design, but this is how many were available in this polygon
                          approximate_segment_count <- area / (160 / 2.47)
                          
                          # This is the sum of the relative weights of the OBSERVED points
                          # This does not include the inaccessible, rejected, or unknown points
                          # It does include both AIM and LMF, however
                          sum_observed_relwgts <- sum(points[points[["observed"]], "relwgt"])
                          # This is the sum of the relative weights of all the AIM points, regardless of fate
                          sum_relwgts <- sum(points[["relwgt"]])
                          
                          # The units are segments per point
                          # The segments are 160 acre quads (quarter sections?) that the first stage of LMF picked from
                          # Steve Garman called this "ConWgt" which I've expanded to conditional_weight
                          # but that's just a guess at what "con" was short for
                          conditional_weight <- approximate_segment_count / (lmf_segment_count)
                          conditional_area <- conditional_weight * sum_observed_relwgts
                          
                          # What's the observed proportion of the area?
                          observed_proportion <- sum_observed_relwgts / sum_relwgts
                          # And how many acres is that then?
                          # We can derive the "unknown" or "unsampled" area as the difference
                          # between the polygon area and the observed area
                          observed_area <- area * observed_proportion
                          
                          # Then this adjustment value is calculated
                          weight_adjustment <- observed_area / conditional_area
                
                          
                          
                          # Put everything about the wgtcat in general in one output data frame
                          output_wgtcat <- data.frame(wgtcat = X,
                                                      area = area,
                                                      area_units = "hectares",
                                                      approximate_segment_count = approximate_segment_count,
                                                      sum_observed_relwgts = sum_observed_relwgts,
                                                      sum_relwgts = sum_relwgts,
                                                      observed_proportion = observed_proportion,
                                                      observed_area = observed_area,
                                                      unobserved_area = area - observed_area,
                                                      conditional_weight = conditional_weight,
                                                      conditional_area = conditional_area,
                                                      weight_adjustment = weight_adjustment,
                                                      point_count = sum(points[["observed"]]),
                                                      observed_point_count = nrow(points),
                                                      stringsAsFactors = FALSE)
                          
                          # But much more importantly add the calculated weights to the points
                          output_points <- points
                          
                          # This handles situations where there were points, but none of them were observed
                          # In that case, weight_adjustment will be 0 / 0 = NaN
                          # That's fine because we can identify that this polygon is still in the inference area
                          # but that its entire area is "unknown" because no data were in it
                          if (is.nan(weight_adjustment)) {
                            output_points[["wgt"]] <- NA
                          } else {
                            # The point weight is calculated here.
                            # This is Garman's formula and I don't have the documentation justifying it on hand
                            output_points[["wgt"]] <- weight_adjustment * conditional_weight * points[["relwgt"]]
                            message("Checking weight sum for ", X)
                            # I'm rounding here because at unrealistically high levels of precision it gets weird and can give false positives
                            if (round(sum(output_points[["wgt"]]), digits = 3) != round(area, digits = 3)) {
                              warning("The sum of the point weights (", sum(output_points[["wgt"]]), ") does not equal the polygon area (", area, ") for ", X)
                            }
                          }
                          
                        } else {
                          # Basically just empty data frames
                          output_wgtcat <- data.frame(wgtcat = X,
                                                      area = area,
                                                      area_units = "hectares",
                                                      approximate_segment_count = area / (160 / 2.47),
                                                      sum_observed_relwgts = NA,
                                                      sum_relwgts = NA,
                                                      observed_proportion = 0,
                                                      observed_area = 0,
                                                      unobserved_area = area,
                                                      conditional_weight = NA,
                                                      conditional_area = NA,
                                                      weight_adjustment = NA,
                                                      point_count = 0,
                                                      observed_point_count = 0,
                                                      stringsAsFactors = FALSE)
                          output_points <- NULL
                        }
                        
                        return(list(points = output_points,
                                    wgtcat = output_wgtcat))
                      })

point_weights <- do.call(rbind,
                         lapply(X = weight_info,
                                FUN = function(X){
                                  X[["points"]]
                                }))
wgtcat_summary <- do.call(rbind,
                          lapply(X = weight_info,
                                 FUN = function(X){
                                   X[["wgtcat"]]
                                 }))

```

```{r}
# Lets join lmf points to weights for output
lmf_points <- sp::merge(x = lmf_points,
                        y = point_weights[,c("unique_id","aim","lmf","observed","relwgt","wgt")],
                        by = "unique_id")
```


### Writing Outputs
```{r}
write.csv(point_weights,
          file = paste0(output_path, "/",
                        output_filename,
                        "_pointweights_",
                        date, ".csv"),
          row.names = FALSE)
write.csv(wgtcat_summary,
          file = paste0(output_path, "/",
                        output_filename,
                        "_wgtcatsummary_",
                        date, ".csv"),
          row.names = FALSE)

# should write these to GDB to see if that solves weird geometry issues
arc.write(data = lmf_points,
          path = paste0(output_path,"/",
                        output_gdb,"/",output_filename,
                               "_points"),
          overwrite = TRUE,
          validate = TRUE)

wgtcat_spdf <- sp::merge(x = wgtcat_spdf,
                         y = wgtcat_summary,
                         by = "wgtcat")

arc.write(data = wgtcat_spdf,
          path = paste0(output_path,"/",
                        output_gdb,"/",output_filename,
                               "_wgtcats"),
          overwrite = TRUE,
          validate = TRUE) 
```


