---
title: "Landscape Monitoring Framework Estimates By State By Type: BLM Rangelands West-Wide"
date: "4/16/2020"
output: word_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 

```{r prep the data, echo = FALSE, include = FALSE}
## Sage Grouse Indicators by State by Type

library(tidyverse)
library(ggplot2)

#this file has 2012 values replaced with year 2013 since the surveys actually took place in 2013
StateType <- read.csv(file = "C:/Users/rburke/Documents/Projects/LMF_DataWrangling/GetTidy/EstimatesBLMByType_2012_2019_withlabels.csv", stringsAsFactors = FALSE)


StateType[["Type"]][is.na(StateType[["Type"]])] <- "Overall"
StateType$Type <- gsub("1" , "I", StateType$Type)
StateType$Type <- gsub("2" , "II", StateType$Type)
#R is not reading numbers as numeric...fixing
#StateType_na <- StateType %>% filter(is.na(Value))
#corrected_strings <- gsub(StateType[["Value"]], pattern = "\\D", replacement = ".")


StateType_Subset <- StateType %>% mutate(Plus = CI_Upper - Value, Minus = Value - CI_Lower) 

write.csv(StateType_Subset, file = "Estimates_BLMByType_WithTableValues.csv", row.names = FALSE)

# Replace 2012 with 2013 because the sampling was actually in 2013
StateType$Year <- gsub("2012" , "2013", StateType$Year)
# StateType$Year <- as.numeric(as.character(StateType$Year))

#Subset to state of interest
#StateType <- StateType %>% filter(!is.na(Type)) %>% subset(State == "WY")

Type_Names <- c("1" = "Type I" , "2" = "Type II", "Overall" = "Overall")

str(StateType)

StateType$Value <- as.numeric(as.character(StateType$Value))
StateType$SE <- as.numeric(as.character(StateType$SE))
StateType$CV <- as.numeric(as.character(StateType$CV))
StateType$CI_Lower <- as.numeric(as.character(StateType$CI_Lower))
StateType$CI_Upper <- as.numeric(as.character(StateType$CI_Upper))

#Pull the max value of acres for plot scale
#Put acres in millions

StateType <- mutate(StateType, Value2 = ifelse(var_type == "acres", Value/1000000, Value)) %>% 
             mutate(CI_Lower2 = ifelse(var_type == "acres", CI_Lower/1000000, CI_Lower)) %>% 
             mutate(CI_Upper2 = ifelse(var_type == "acres", CI_Upper/1000000, CI_Upper))


Max_acres <- StateType %>% subset(var_type == "acres") %>% summarize(max(CI_Upper2)) %>% round(digits = 2)

Max_acres <- Max_acres$`max(CI_Upper2)` 


#Max_inches <- StateType %>% subset(var_type == "inches") %>% summarize(max(CI_Upper2)) %>% round(digits = 2)

#Max_inches <- Max_inches$`max(CI_Upper2)` + 0.5

#Grabbing Max Inches from other doc instead to keep consistent

Max_inches <- 68.16

#StateType[["Value"]] <- trimws(StateType[["Value"]])


# Split the title into 2 lines for labelling

#StateType <- StateType %>% mutate(figure_title_caption = figure_title)

StateName <- c("AZ" = "Arizona", "CA" = "California", "CO" = "Colorado" , "ID" = "Idaho" , 
               "MT" = "Montana", "ND" = "North Dakota", "NM" = "New Mexico" , "NV" = "Nevada",
               "OR" = "Oregon", "SD" = "South Dakota", "UT" = "Utah", "WA" = "Washington", 
               "WY" = "Wyoming", "BLM" = "BLM Rangelands West-Wide")

str(StateType)
StateType$Year <- as.numeric(as.character(StateType$Year))

#Remove overall estimate and remove 2011 because it's not available across the dataset
StateType_TypeOnly <- StateType %>% subset(Type == "I" | Type == "II") %>% filter(!Year == "2011")
unique(StateType_TypeOnly$Indicator_Name) %>% unique()

StateType_TypeOnly_CIFix <- StateType_TypeOnly %>% filter(Year == "2013" & Indicator_Name == "67. Mean Sagebrush Species Height in Inches on BLM range")

list_of_plots <- lapply(X = split(StateType_TypeOnly, StateType_TypeOnly[["Index"]]), FUN = function(StateType_TypeOnly){

  current_plot <- ggplot(StateType_TypeOnly, aes(x = Year, y = Value2, fill = Type, group = Type)) +
                  geom_point(aes(x = Year , 
                  y = Value2, shape = Type, fill = Type), position = position_dodge2(.5), size = 2) + 
                  labs(x = StateType_TypeOnly$xlab , y = StateType_TypeOnly$ylab, fig.caption = StateType_TypeOnly$caption) +
                  theme(plot.caption = element_text(hjust=0.5, size=rel(1.2), face = "bold.italic")) + 
                  scale_fill_discrete(name = "Type", labels = c("I", "II")) + 
                  scale_x_continuous(breaks = scales::pretty_breaks(n = length(unique(StateType_TypeOnly$Year + 1)))) + 
                  geom_errorbar(aes(x = Year , ymin = CI_Lower2, ymax = CI_Upper2, fill = Type) , width = 0.2, position = position_dodge(.5)) +  
                  scale_shape_manual(values = c(16,17)) + 
                  guides(size = "none", color = "none", fill = "none") + 
                  facet_grid(cols = vars(State), labeller = labeller(State = StateName)) + 
                  theme(strip.text.x = element_text(size = 10 , color = "black" , face = "bold") ,
                  plot.caption = element_text(face = "italic", size = "8", hjust = 0), 
                  plot.subtitle = element_text(face = "italic", size = "8", hjust = 0), 
                  panel.background = element_rect(fill = "white") , panel.grid = element_line(color = "grey") ,
                  panel.grid.minor = element_blank() , panel.grid.major.x = element_line(color = "grey95")) +
    
                  if(StateType_TypeOnly$var_type == "percent"){scale_y_continuous(limits = c(0 , 100))} 
                  else if(StateType_TypeOnly$var_type == "acres"){scale_y_continuous(limits = c(0, Max_acres))} 
                  else if(StateType_TypeOnly$var_type == "inches"){scale_y_continuous(limits = c(0 , Max_inches))}
   
   return(current_plot)
  
})

    # if(StateType_TypeOnly$Indicator_Number.x < "27"){scale_x_continuous(breaks = scales::pretty_breaks(n = 8))}
    # else if(StateType_TypeOnly$Indicator_Number.x >= "27"){scale_x_continuous(breaks = scales::pretty_breaks(n = 6))}

## Removing title for formatting issues
# Will embed in doc
# title =  paste("Figure ", StateType$Index, ":", StateType$figure_title), face = "italic"

```

#### ***Figure 1: Proportion of BLM rangelands with none-to-slight or slight-to-moderate departure from biotic integrity reference conditions (BI <=2) (80% confidence interval)*** 

```{r fig 1, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["1"]]

```


#### ***Figure 2: BLM rangeland acres with none-to-slight or slight-to-moderate departure from biotic integrity reference conditions (BI <=2) (80% confidence interval)*** 

```{r fig 2, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["2"]]

```

#### ***Figure 3: Proportion of BLM rangelands with moderate departure from biotic integrity reference conditions (BI = 3) (80% confidence interval)*** 

```{r fig 3, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["3"]]

```

#### ***Figure 4: BLM rangeland acres with moderate departure from biotic integrity reference conditions (BI = 3) (80% confidence interval)*** 

```{r fig 4, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["4"]]

```

#### ***Figure 5: Proportion of BLM rangelands with moderate or moderate-to-extreme or extreme-to-total departure from biotic integrity reference conditions (BI >=3) (80% confidence interval)*** 

```{r fig 5, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["5"]]

```

#### ***Figure 6: BLM rangeland acres with moderate or moderate-to-extreme or extreme-to-total departure from biotic integrity reference conditions (BI >=3) (80% confidence interval)*** 

```{r fig 6, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["6"]]

```

#### ***Figure 7: Proportion of BLM rangelands with moderate-to-extreme or extreme-to-total departure from biotic integrity reference conditions (BI >= 4) (80% confidence interval)*** 

```{r fig 7, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["7"]]

```

#### ***Figure 8: BLM rangeland acres with moderate-to-extreme or extreme-to-total departure from biotic integrity reference conditions (BI >= 4) (80% confidence interval)*** 

```{r fig 8, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["8"]]

```

#### ***Figure 9: Proportion of BLM rangelands with none-to-slight or slight-to-moderate departure from hydrologic function reference conditions (HF <=2) (80% confidence interval)*** 

```{r fig 9, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["9"]]

```

#### ***Figure 10: BLM rangeland acres with none-to-slight or slight-to-moderate departure from hydrologic function reference conditions (HF <=2) (80% confidence interval)*** 

```{r fig 10, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["10"]]

```

#### ***Figure 11: Proportion of BLM rangelands with moderate departure from hydrologic function reference conditions (HF = 3) (80% confidence interval)*** 

```{r fig 11, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["11"]]

```

#### ***Figure 12: BLM rangeland acres with moderate departure from hydrologic function reference conditions (HF = 3) (80% confidence interval)*** 

```{r fig 12, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["12"]]

```

#### ***Figure 13: Proportion of BLM rangelands with moderate or moderate-to-extreme or extreme-to-total departure from hydrologic function reference conditions (HF >= 3) (80% confidence interval)*** 

```{r fig 13, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["13"]]

```

#### ***Figure 14: BLM rangeland acres with moderate or moderate-to-extreme or extreme-to-total departure from hydrologic function reference conditions (HF >= 3) (80% confidence interval)*** 

```{r fig 14, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["14"]]

```

#### ***Figure 15: Proportion of BLM rangelands with moderate-to-extreme or extreme-to-total departure from hydrologic function reference conditions (HF >= 4) (80% confidence interval)*** 

```{r fig 15, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["15"]]

```

#### ***Figure 16: BLM rangeland acres with moderate-to-extreme or extreme-to-total departure from hydrologic function reference conditions (HF >= 4) (80% confidence interval)*** 

```{r fig 16, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["16"]]

```

#### ***Figure 17: Proportion of BLM rangelands with none-to-slight or slight-to-moderate departure from soil and site stability reference conditions (SSS <=2) (80% confidence interval) *** 

```{r fig 17, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["17"]]

```

#### ***Figure 18: BLM rangeland acres with none-to-slight or slight-to-moderate departure from soil and site stability reference conditions (SSS <=2) (80% confidence interval)*** 

```{r fig 18, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["18"]]

```

#### ***Figure 19: Proportion of BLM rangelands with moderate departure from soil and site stability reference conditions (SSS = 3) (80% confidence interval)*** 

```{r fig 19, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["19"]]

```

#### ***Figure 20: BLM rangeland acres with moderate departure from soil and site stability reference conditions (SSS = 3) (80% confidence interval)*** 

```{r fig 20, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["20"]]

```

#### ***Figure 21: Proportion of BLM rangelands with moderate or moderate-to-extreme or extreme-to-total departure from soil and site stability reference conditions (SSS >= 3) (80% confidence interval)*** 

```{r fig 21, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["21"]]

```

#### ***Figure 22: BLM rangeland acres with moderate or moderate-to-extreme or extreme-to-total departure from soil and site stability reference conditions (SSS >= 3) (80% confidence interval)*** 

```{r fig 22, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["22"]]

```

#### ***Figure 23: Proportion of BLM rangelands with moderate-to-extreme or extreme-to-total departure from soil and site stability reference conditions (SSS >= 4) (80% confidence interval)*** 

```{r fig 23, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["23"]]

```

#### ***Figure 24: BLM rangeland acres with moderate-to-extreme or extreme-to-total departure from soil and site stability reference conditions (SSS >= 4) (80% confidence interval)*** 

```{r fig 24, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["24"]]

```

#### ***Figure 25: Proportion of BLM rangelands where all 3 rangeland health attributes had none-to-slight or slight-to-moderate departure from reference conditions (<= 2) (80% confidence interval)*** 

```{r fig 25, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["25"]]

```

#### ***Figure 26: BLM rangeland acres where all 3 rangeland health attributes had none-to-slight or slight-to-moderate departure from reference conditions (<= 2) (80% confidence interval)*** 

```{r fig 26, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["26"]]

```

#### ***Figure 27: Proportion of BLM rangelands where all 3 rangeland health attributes had moderate-to-extreme or extreme-to-total departure from reference conditions (>= 4) (80% confidence interval)*** 

```{r fig 27, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["27"]]

```

#### ***Figure 28: BLM rangeland acres where all 3 rangeland health attributes had moderate-to-extreme or extreme-to-total departure from reference conditions (>= 4) (80% confidence interval)*** 

```{r fig 28, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["28"]]

```

#### ***Figure 29: Proportion of BLM rangelands where at least 1 rangeland health attribute had moderate departure from reference conditions (= 3) (80% confidence interval)*** 

```{r fig 29, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["29"]]

```

#### ***Figure 30: BLM rangeland acres where at least 1 rangeland health attribute had moderate departure from reference conditions (= 3) (80% confidence interval)*** 

```{r fig 30, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["30"]]

```

#### ***Figure 31: Proportion of BLM rangelands with non-native invasive species present (80% confidence interval)*** 

```{r fig 31, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["31"]]

```

#### ***Figure 32: BLM rangeland acres with non-native invasive species present (80% confidence interval)*** 

```{r fig 32, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["32"]]

```

#### ***Figure 33: Proportion of BLM rangelands where non-native invasive species are abundant (>= 25%) (80% confidence interval)*** 

```{r fig 33, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["33"]]

```

#### ***Figure 34: BLM rangeland acres where non-native invasive species are abundant (>= 25%) (80% confidence interval)*** 

```{r fig 34, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["34"]]

```

#### ***Figure 35: Proportion of BLM rangelands with at least 20% of land having intercanopy gaps of at least 2 meters (80% confidence interval)*** 

```{r fig 35, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["35"]]

```

#### ***Figure 36: BLM rangeland acres with at least 20% of land having intercanopy gaps of at least 2 meters in length (80% confidence interval)*** 

```{r fig 36, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["36"]]

```

#### ***Figure 37: Proportion of BLM rangelands with at least 20% of land having intercanopy gaps of at least 2 meters and gaps are at least 50% bare ground (80% confidence interval)*** 

```{r fig 37, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["37"]]

```

#### ***Figure 38: BLM rangeland acres with at least 20% of land having intercanopy gaps of at least 2 meters and gaps are at least 50% bare ground (80% confidence interval)*** 

```{r fig 38, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["38"]]

```

#### ***Figure 39: Percent cover of bare ground on BLM rangelands (80% confidence interval)*** 

```{r fig 39, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["39"]]

```

#### ***Figure 40: Proportion of BLM rangelands with at least 20% bare ground (80% confidence interval)*** 

```{r fig 40, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["40"]]

```

#### ***Figure 41: BLM rangeland acres with at least 20% bare ground (80% confidence interval)*** 

```{r fig 41, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["41"]]

```

#### ***Figure 42: Proportion of BLM rangelands with at least 30% bare ground (80% confidence interval)*** 

```{r fig 42, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["42"]]

```

#### ***Figure 43: BLM rangeland acres with at least 30% bare ground (80% confidence interval)*** 

```{r fig 43, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["43"]]
```

#### ***Figure 44: Proportion of BLM rangelands with at least 40% bare ground (80% confidence interval)*** 

```{r fig 44, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["44"]]

```

#### ***Figure 45: BLM rangeland acres with at least 40% bare ground (80% confidence interval)*** 

```{r fig 45, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["45"]]

```

#### ***Figure 46: Proportion of BLM rangelands with at least 50% bare ground (80% confidence interval)*** 

```{r fig 46, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["46"]]

```

#### ***Figure 47: BLM rangeland acres with at least 50% bare ground (80% confidence interval)*** 

```{r fig 47, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["47"]]

```

#### ***Figure 48: Proportion of BLM rangelands where soil aggregate stability is ranked 4 or less (80% confidence interval)*** 

```{r fig 48, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["48"]]

```

#### ***Figure 49: BLM rangeland acres where soil aggregate stability is ranked 4 or less (80% confidence interval)*** 

```{r fig 49, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["49"]]

```

#### ***Figure 50: Percent cover of sagebrush on BLM rangelands (80% confidence interval)*** 

```{r fig 50, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["50"]]

```

#### ***Figure 51: Mean sagebrush species height in inches on BLM rangelands (80% confidence interval)*** 

```{r fig 51, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["51"]]

```

#### ***Figure 52: Proportion of sagebrush that is columnar shaped on BLM rangelands (80% confidence interval)*** 

```{r fig 52, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["52"]]

```

#### ***Figure 53: Proportion of sagebrush that is spreading shaped on BLM rangelands (80% confidence interval)*** 

```{r fig 53, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["53"]]

```

#### ***Figure 54: Mean herbaceous plant species height in inches on BLM rangelands (80% confidence interval)*** 

```{r fig 54, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["54"]]

```

#### ***Figure 55: Percent cover of perennial grasses and perennial forbs on BLM rangelands (80% confidence interval)*** 

```{r fig 55, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["55"]]

```

#### ***Figure 56: Percent cover of bare ground on BLM rangelands (80% confidence interval)*** 

```{r fig 56, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["56"]]

```

#### ***Figure 57: Proportion of vegetation composed of annual grasses on BLM rangelands (80% confidence interval)*** 

```{r fig 57, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["57"]]

```

#### ***Figure 58: Proportion of vegetation composed of non-native invasive plant species on BLM rangelands (80% confidence interval)*** 

```{r fig 58, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["58"]]

```



