---
title: "Landscape Monitoring Framework Estimates By State By Type: Montana"
date: "10/6/2020"
output: word_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 

```{r prep the data, echo = FALSE, include = FALSE}
## Sage Grouse Indicators by State by Type

library(tidyverse)
library(ggplot2)

#this file has 2012 values replaced with year 2013 since the surveys actually took place in 2013
StateType <- read.csv(file = "EstimatesByStateByType_2012_2019_withlabels.csv", stringsAsFactors = FALSE)
unique(StateType$State)
str(StateType)

#R is not reading numbers as numeric...fixing in excel because I cant figure it out

# Replace 2012 with 2013 because the sampling was actually in 2013
StateType$Year <- gsub("2012" , "2013", StateType$Year)
StateType$Year <- as.integer(as.character(StateType$Year))
#Read in sample size and remove estimates for years with fewer than 10 samples

Sample <- read.csv("SampleSizeByStateType_Year.csv", stringsAsFactors = FALSE)

## Why is COlorado sample size for 2019 disappearing 
colorado2019 <- Sample %>% subset(State == "MT")
trimws(Sample$State)
#Merge sample with the whole dataframe
StateType <- full_join(StateType, Sample, by = c("State", "Year", "Type"))
StateTypesubset <- StateType %>% subset(State == "MT" & Year == "2019")

StateType <- StateType %>% mutate(caption = ifelse(BLM.Sample < 10 , paste0(caption, "* Missing values: Estimates unavailable due to small sample size (fewer than 10 plots)", sep = " "), caption))

# Join with axis label doc
#I already did this in the GetTidy code
# Labels <- read.csv(file = "FigureLabels_updated.csv")
# StateType <- merge(StateType, Labels, by = "Indicator_Name") 
# StateType <- StateType %>% select(-Indicator.y, -Indicator_Number.y, -Indicator.x)


#Subset to state of interest
StateType <- StateType %>% filter(!is.na(Type)) %>% subset(State == "MT")
#For some reason we got duplicates in here somehow
StateType <- StateType %>% unique()

Type_Names <- c("1" = "Type I" , "2" = "Type II")

#Pull the max value of acres for plot scale
#Put acres in millions

StateType <- mutate(StateType, Value2 = ifelse(var_type == "acres", Value/1000000, Value)) %>% 
             mutate(CI_Lower2 = ifelse(var_type == "acres", CI_Lower/1000000, CI_Lower)) %>% 
             mutate(CI_Upper2 = ifelse(var_type == "acres", CI_Upper/1000000, CI_Upper))


Max_acres <- StateType %>% subset(var_type == "acres") %>% summarize(max(Value2)) %>% round(digits = 2)
Max_acres_CI <- StateType %>% subset(var_type == "acres") %>% summarize(max(CI_Upper2)) %>% round(digits = 2)
Max_acres <- (Max_acres$`max(Value2)`) + (Max_acres_CI$`max(CI_Upper2)`)

#MAke Value2 NA for samples less than 10
StateType <- StateType %>% mutate(Value2 = ifelse(BLM.Sample < 10 , NA, Value2)) %>% 
                           mutate(CI_Upper2 = ifelse(BLM.Sample < 10 , NA, CI_Upper2)) %>%
                           mutate(CI_Lower2 = ifelse(BLM.Sample < 10 , NA, CI_Lower2))


#StateType[["Value"]] <- trimws(StateType[["Value"]])
str(StateType)

StateType$Value <- as.numeric(as.character(StateType$Value))
StateType$SE <- as.numeric(as.character(StateType$SE))
StateType$CV <- as.numeric(as.character(StateType$CV))
StateType$CI_Lower <- as.numeric(as.character(StateType$CI_Lower))
StateType$CI_Upper <- as.numeric(as.character(StateType$CI_Upper))

# Split the title into 2 lines for labelling

StateType <- StateType %>% mutate(figure_title_caption = figure_title)

StateName <- c("AZ" = "Arizona", "CA" = "California", "CO" = "Colorado" , "ID" = "Idaho" , 
               "MT" = "Montana", "ND" = "North Dakota", "NM" = "New Mexico" , "NV" = "Nevada",
               "OR" = "Oregon", "SD" = "South Dakota", "UT" = "Utah", "WA" = "Washington", 
               "WY" = "Wyoming")

StateType$Type <- as.factor(as.integer(StateType$Type))

StateType <- StateType %>% mutate_if(is.numeric, round, digits = 2)

list_of_plots <- lapply(X = split(StateType, StateType[["Index"]]), FUN = function(StateType){

  current_plot <- ggplot(StateType, aes(x = Year, y = Value2, fill = Type, group = Type)) +
                  geom_point(aes(x = Year , 
                  y = Value2, shape = Type, fill = Type), position = position_dodge2(1), size = 2) + 
    labs(x = StateType$xlab , y = StateType$ylab2, caption = StateType$caption) +
    theme(plot.caption = element_text(hjust=0, size=rel(1), face = "bold.italic")) + 
    scale_fill_discrete(name = "Type", labels = c("I", "II")) + 
    geom_errorbar(aes(x = Year , ymin = CI_Lower2 ,
                      ymax = CI_Upper2, fill = Type) , width = 0.2, position = position_dodge(1)) +  
    scale_x_continuous(breaks = scales::pretty_breaks(n = length(unique(StateType$Year)))) +
    scale_shape_manual(values = c(16,17)) + 
    guides(size = "none", color = "none", fill = "none") + 
    facet_grid(cols = vars(State), labeller = labeller(State = StateName)) + 
    theme(strip.text.x = element_text(size = 10 , color = "black" , face = "bold") ,
          panel.background = element_rect(fill = "white") , panel.grid = element_line(color = "grey") ,
          panel.grid.minor = element_blank() , panel.grid.major.x = element_line(color = "grey95")) +
  
    if(StateType$var_type == "percent"){scale_y_continuous(limits = c(0 , 100))} 
    else if(StateType$var_type == "acres"){scale_y_continuous(limits = c(0, Max_acres))} 
    else if(StateType$var_type == "inches"){scale_y_continuous(limits = c(0 , 50))}

    return(current_plot)
  
})

## Removing title for formatting issues
# Will embed in doc
# title =  paste("Figure ", StateType$Index, ":", StateType$figure_title), face = "italic"

```

#### ***Figure 1: Proportion of BLM rangelands with none-to-slight or slight-to-moderate departure from biotic integrity reference conditions (BI <=2) (80% confidence interval)*** 

```{r fig 1, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["1"]]

```


#### ***Figure 2: BLM rangeland acres with none-to-slight or slight-to-moderate departure from biotic integrity reference conditions (BI <=2) (80% confidence interval)*** 

```{r fig 2, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["2"]]

```

#### ***Figure 3: Proportion of BLM rangelands with moderate departure from biotic integrity reference conditions (BI = 3) (80% confidence interval)*** 

```{r fig 3, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["3"]]

```

#### ***Figure 4: BLM rangeland acres with moderate departure from biotic integrity reference conditions (BI = 3) (80% confidence interval)*** 

```{r fig 4, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["4"]]

```

#### ***Figure 5: Proportion of BLM rangelands with moderate or moderate-to-extreme or extreme-to-total departure from biotic integrity reference conditions (BI >=3) (80% confidence interval)*** 

```{r fig 5, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["5"]]

```

#### ***Figure 6: BLM rangeland acres with moderate or moderate-to-extreme or extreme-to-total departure from biotic integrity reference conditions (BI >=3) (80% confidence interval)*** 

```{r fig 6, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["6"]]

```

#### ***Figure 7: Proportion of BLM rangelands with moderate-to-extreme or extreme-to-total departure from biotic integrity reference conditions (BI >= 4) (80% confidence interval)*** 

```{r fig 7, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["7"]]

```

#### ***Figure 8: BLM rangeland acres with moderate-to-extreme or extreme-to-total departure from biotic integrity reference conditions (BI >= 4) (80% confidence interval)*** 

```{r fig 8, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["8"]]

```

#### ***Figure 9: Proportion of BLM rangelands with none-to-slight or slight-to-moderate departure from hydrologic function reference conditions (HF <=2) (80% confidence interval)*** 

```{r fig 9, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["9"]]

```

#### ***Figure 10: BLM rangeland acres with none-to-slight or slight-to-moderate departure from hydrologic function reference conditions (HF <=2) (80% confidence interval)*** 

```{r fig 10, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["10"]]

```

#### ***Figure 11: Proportion of BLM rangelands with moderate departure from hydrologic function reference conditions (HF = 3) (80% confidence interval)*** 

```{r fig 11, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["11"]]

```

#### ***Figure 12: BLM rangeland acres with moderate departure from hydrologic function reference conditions (HF = 3) (80% confidence interval)*** 

```{r fig 12, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["12"]]

```

#### ***Figure 13: Proportion of BLM rangelands with moderate or moderate-to-extreme or extreme-to-total departure from hydrologic function reference conditions (HF >= 3) (80% confidence interval)*** 

```{r fig 13, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["13"]]

```

#### ***Figure 14: BLM rangeland acres with moderate or moderate-to-extreme or extreme-to-total departure from hydrologic function reference conditions (HF >= 3) (80% confidence interval)*** 

```{r fig 14, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["14"]]

```

#### ***Figure 15: Proportion of BLM rangelands with moderate-to-extreme or extreme-to-total departure from hydrologic function reference conditions (HF >= 4) (80% confidence interval)*** 

```{r fig 15, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["15"]]

```

#### ***Figure 16: BLM rangeland acres with moderate-to-extreme or extreme-to-total departure from hydrologic function reference conditions (HF >= 4) (80% confidence interval)*** 

```{r fig 16, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["16"]]

```

#### ***Figure 17: Proportion of BLM rangelands with none-to-slight or slight-to-moderate departure from soil and site stability reference conditions (SSS <=2) (80% confidence interval) *** 

```{r fig 17, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["17"]]

```

#### ***Figure 18: BLM rangeland acres with none-to-slight or slight-to-moderate departure from soil and site stability reference conditions (SSS <=2) (80% confidence interval)*** 

```{r fig 18, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["18"]]

```

#### ***Figure 19: Proportion of BLM rangelands with moderate departure from soil and site stability reference conditions (SSS = 3) (80% confidence interval)*** 

```{r fig 19, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["19"]]

```

#### ***Figure 20: BLM rangeland acres with moderate departure from soil and site stability reference conditions (SSS = 3) (80% confidence interval)*** 

```{r fig 20, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["20"]]

```

#### ***Figure 21: Proportion of BLM rangelands with moderate or moderate-to-extreme or extreme-to-total departure from soil and site stability reference conditions (SSS >= 3) (80% confidence interval)*** 

```{r fig 21, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["21"]]

```

#### ***Figure 22: BLM rangeland acres with moderate or moderate-to-extreme or extreme-to-total departure from soil and site stability reference conditions (SSS >= 3) (80% confidence interval)*** 

```{r fig 22, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["22"]]

```

#### ***Figure 23: Proportion of BLM rangelands with moderate-to-extreme or extreme-to-total departure from soil and site stability reference conditions (SSS >= 4) (80% confidence interval)*** 

```{r fig 23, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["23"]]

```

#### ***Figure 24: BLM rangeland acres with moderate-to-extreme or extreme-to-total departure from soil and site stability reference conditions (SSS >= 4) (80% confidence interval)*** 

```{r fig 24, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["24"]]

```

#### ***Figure 25: Proportion of BLM rangelands where all 3 rangeland health attributes had none-to-slight or slight-to-moderate departure from reference conditions (<= 2) (80% confidence interval)*** 

```{r fig 25, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["25"]]

```

#### ***Figure 26: BLM rangeland acres where all 3 rangeland health attributes had none-to-slight or slight-to-moderate departure from reference conditions (<= 2) (80% confidence interval)*** 

```{r fig 26, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["26"]]

```

#### ***Figure 27: Proportion of BLM rangelands where all 3 rangeland health attributes had moderate-to-extreme or extreme-to-total departure from reference conditions (>= 4) (80% confidence interval)*** 

```{r fig 27, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["27"]]

```

#### ***Figure 28: BLM rangeland acres where all 3 rangeland health attributes had moderate-to-extreme or extreme-to-total departure from reference conditions (>= 4) (80% confidence interval)*** 

```{r fig 28, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["28"]]

```

#### ***Figure 29: Proportion of BLM rangelands where at least 1 rangeland health attribute had moderate departure from reference conditions (= 3) (80% confidence interval)*** 

```{r fig 29, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["29"]]

```

#### ***Figure 30: BLM rangeland acres where at least 1 rangeland health attribute had moderate departure from reference conditions (= 3) (80% confidence interval)*** 

```{r fig 30, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["30"]]

```

#### ***Figure 31: Proportion of BLM rangelands with non-native invasive species present (80% confidence interval)*** 

```{r fig 31, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["31"]]

```

#### ***Figure 32: BLM rangeland acres with non-native invasive species present (80% confidence interval)*** 

```{r fig 32, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["32"]]

```

#### ***Figure 33: Proportion of BLM rangelands where non-native invasive species are abundant (>= 25%) (80% confidence interval)*** 

```{r fig 33, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["33"]]

```

#### ***Figure 34: BLM rangeland acres where non-native invasive species are abundant (>= 25%) (80% confidence interval)*** 

```{r fig 34, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["34"]]

```

#### ***Figure 35: Proportion of BLM rangelands with at least 20% of land having intercanopy gaps of at least 2 meters (80% confidence interval)*** 

```{r fig 35, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["35"]]

```

#### ***Figure 36: BLM rangeland acres with at least 20% of land having intercanopy gaps of at least 2 meters in length (80% confidence interval)*** 

```{r fig 36, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["36"]]

```

#### ***Figure 37: Proportion of BLM rangelands with at least 20% of land having intercanopy gaps of at least 2 meters and gaps are at least 50% bare ground (80% confidence interval)*** 

```{r fig 37, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["37"]]

```

#### ***Figure 38: BLM rangeland acres with at least 20% of land having intercanopy gaps of at least 2 meters and gaps are at least 50% bare ground (80% confidence interval)*** 

```{r fig 38, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["38"]]

```

#### ***Figure 39: Average amount of bare ground on BLM rangelands (80% confidence interval)*** 

```{r fig 39, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["39"]]

```

#### ***Figure 40: Proportion of BLM rangelands with at least 20% bare ground (80% confidence interval)*** 

```{r fig 40, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["40"]]

```

#### ***Figure 41: BLM rangeland acres with at least 20% bare ground (80% confidence interval)*** 

```{r fig 41, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["41"]]

```

#### ***Figure 42: Proportion of BLM rangelands with at least 30% bare ground (80% confidence interval)*** 

```{r fig 42, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["42"]]

```

#### ***Figure 43: BLM rangeland acres with at least 30% bare ground (80% confidence interval)*** 

```{r fig 43, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["43"]]
```

#### ***Figure 44: Proportion of BLM rangelands with at least 40% bare ground (80% confidence interval)*** 

```{r fig 44, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["44"]]

```

#### ***Figure 45: BLM rangeland acres with at least 40% bare ground (80% confidence interval)*** 

```{r fig 45, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["45"]]

```

#### ***Figure 46: Proportion of BLM rangelands with at least 50% bare ground (80% confidence interval)*** 

```{r fig 46, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["46"]]

```

#### ***Figure 47: BLM rangeland acres with at least 50% bare ground (80% confidence interval)*** 

```{r fig 47, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["47"]]

```

#### ***Figure 48: Proportion of BLM rangelands where soil aggregate stability is ranked 4 or less (80% confidence interval)*** 

```{r fig 48, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["48"]]

```

#### ***Figure 49: BLM rangeland acres where soil aggregate stability is ranked 4 or less (80% confidence interval)*** 

```{r fig 49, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["49"]]

```

#### ***Figure 50: Average foliar cover of sagebrush on BLM rangelands (80% confidence interval)*** 

```{r fig 50, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["50"]]

```

#### ***Figure 51: Mean sagebrush species height in inches on BLM rangelands (80% confidence interval)*** 

```{r fig 51, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["51"]]

```

#### ***Figure 52: Proportion of sagebrush that is columnar shaped on BLM rangelands (80% confidence interval)*** 

```{r fig 52, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["52"]]

```

#### ***Figure 53: Proportion of sagebrush that is spreading shaped on BLM rangelands (80% confidence interval)*** 

```{r fig 53, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["53"]]

```

#### ***Figure 54: Mean herbaceous plant species height in inches on BLM rangelands (80% confidence interval)*** 

```{r fig 54, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["54"]]

```

#### ***Figure 55: Proportion of vegetation composed of perennial grasses and perennial forbs on BLM rangelands (80% confidence interval)*** 

```{r fig 55, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["55"]]

```

#### ***Figure 56: Proportion of land composed of bare ground on BLM rangelands (80% confidence interval)*** 

```{r fig 56, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["56"]]

```

#### ***Figure 57: Proportion of vegetation composed of annual grasses on BLM rangelands (80% confidence interval)*** 

```{r fig 57, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["57"]]

```

#### ***Figure 58: Proportion of vegetation composed of non-native invasive plant species on BLM rangelands (80% confidence interval)*** 

```{r fig 58, echo = FALSE , fig.height = 6 , fig.width = 8}

list_of_plots[["58"]]

```



